\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample

### 2017-10-24 document/XSGRB_sample.tex
Intro work
More work on introduction and conclusions

GRBs are interesting as they provide the opportunity to derive new insights
into a very wide range of topics in astrophysics. In that sense GRB
research is very multidisciplinary. The {\it Swift} satellite \citep{Gehrels09}, 
which was launched in 2004, has made
it possible to harvest the great potential in using GRBs as probes of
galaxies in the intergalactic medium, which was already hinted at by
results from earlier missions.

@@

In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy
was secured using low-resolution spectrographs (typically
R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. For a handful of
very bright afterglows high-resolution (typically R$>$20000) spectra were
secured.

@@

>>>>

Gamma-Ray bursts (GRBs) are interesting as they provide the opportunity to
derive new insights into a very wide range of topics in astrophysics. Examples
range from small scale phenomena relating to the properties of highly
relativistic jets, supernova explosions, the interstellar medium, dust extinction
curves, starburst galaxies, chemical and molecular abundances
the ionization state of the intergalactic medium, intervening
absoption systems to standard candles in cosmology.  In that sense GRB research
is very multidisciplinary. The {\it Swift} satellite \citep{Gehrels09}, which
was launched in 2004, has made it possible to harvest the great potential in
using GRBs as probes of galaxies in the intergalactic medium, which was already
hinted at by results from earlier missions \citep[e.g.][]{Ricker04}.

@@

In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy
was secured using low-resolution spectrographs (typically
R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. Low-resolution 
spectroscopy is powerful for a number of reasons, first of all is the ability 
to secure information for even very faint targets. This allows the 
measurements of a number of important parameters such as redshifts and spectral
slopes. For a handful of
very bright afterglows high-resolution (typically R$>$20000) spectra were
secured and for these events much more information about conditions inside the
host galaxies were extracted 
\citep[e.g.,]{}{Fiore, Thoene07, Prochaska07, Vreejswijk07}.

@@

Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
towards about 10$^5$ QSOs have a handful such systems with
$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
\citep[e.g., five in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, 
reveal such
systems in the majority of cases \citep[][and this work]{Jakobsson2006, Fynbo2009}. 
Whereas DLAs
towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
star-forming galaxies over the full history of cosmic star formation from $z
\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
    Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
spectrum from X-rays to the sub-mm) we can hence characterize the properties of
star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
molecular contents, ISM temperatures, UV-flux densities, etc.

Bias still, more to do, ELTs, SVOM
